---
title: "p8105_hw5_yy3295"
author: "Ye Yuan"
date: "2023-11-15"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

For this problem, we are interested in data gathered and made public by _The Washington Post_ on homicides in 50 large U.S. cities. The code chunk below imports and cleans the data.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) %>% 
  filter(city_state != "Tulsa, AL") 
```

The resulting dataframe has `r nrow(homicide_df)` entries, on variables that include the victim name, race, age, and sex; the date the homicide was reported; and the location of the homicide. In cleaning, I created a `city_state` variable that includes both city and state, and a `resolution` variable to indicate whether the case was closed by arrest. I also excluded one entry in Tulsa, AL, which is not a major US city and is most likely a data entry error. 

In the next code chunk, I group within cities and summarize to produce the total number of homicides and the number that are solved. 

```{r}
city_homicide_df = 
  homicide_df %>% 
  select(city_state, disposition, resolution) %>% 
  group_by(city_state) %>% 
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolution == "unsolved"))
```

Focusing only on Baltimore, MD, I can use the `prop.test` and `broom::tidy` functions to obtain an estimate and CI of the proportion of unsolved homicides in that city. The table below shows those values.

```{r}
bmore_test = 
  prop.test(
    x = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_unsolved),
    n = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_total)) 

broom::tidy(bmore_test) %>% 
  knitr::kable(digits = 3)
```

Building on this code, I can use functions in the `purrr` package to obtain estimates and CIs for the proportion of unsolved homicides in each city in my dataset. The code below implements this analysis. 

```{r}
test_results = 
  city_homicide_df %>% 
  mutate(
    prop_tests = map2(hom_unsolved, hom_total, \(x, y) prop.test(x = x, n = y)),
    tidy_tests = map(prop_tests, broom::tidy)) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high) %>% 
  mutate(city_state = fct_reorder(city_state, estimate))
```

Finally, I make a plot showing the estimate (and CI) of the proportion of unsolved homicides in each city.

```{r}
test_results %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This figure suggests a very wide range in the rate at which homicides are solved -- Chicago is noticeably high and, given the narrowness of the CI, likely is the location of many homicides. 

## Problem 2

```{r}

```


## Problem 3

First set the following design elements.
```{r}
set.seed(1)
```

Simulation: T-test
```{r}
sim_t_test = function(n = 30, mu, sd = 5) {
  x = rnorm(n, mu, sd)
  t.test(x) |>
  broom::tidy() |>
    select(p_value = p.value, mu_hat = estimate)
}
```

Simulation for mean = 0
```{r}
output = vector("list", length = 5000)

for (i in 1:5000) {
  output[[i]] = sim_t_test(mu = 0)
}

tt_results = bind_rows(output)
```

Simulation for mean = 1, 2, 3, 4, 5, 6
```{r}
sim_results_df = 
  expand_grid(
    mus = c(1, 2, 3, 4, 5, 6),
    iter = 1:5000) |>
  mutate(estimate_df = map(mus, ~sim_t_test(n = 30, mu = .))) |>
  unnest(cols = c(estimate_df))
```

Make a plot showing the proportion of times the null was rejected on the y axis and the true value of μ on the x axis. 
Describe the association between effect size and power.
```{r}
plot_q3 = sim_results_df |>
  group_by(mus) |>
  summarise(power = sum(p_value < 0.05) / 5000) |>
  ggplot(
    aes(x = mus,
        y = power)) +
  geom_line() +
  geom_point() +
  labs(x = "True Mean",
       y = "Power",
       title = "True Mean vs Power")

plot_q3
```
From the plot we can learned when the true mean increases, the power increases as well and will eventually reaches 1.00.
Once the effect size increases, the statistical power increases accordingly.

Make a plot showing the average estimate of μ^ on the y axis and the true value of μ on the x axis. 
Make a second plot showing the average estimate of μ^ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. 
Is the sample average of μ^ across tests for which the null is rejected approximately equal to the true value of μ? 
Why or why not?
```{r}
estimate_mu = sim_results_df |>
  group_by(mus) |>
  summarise(average_estimate = mean(mu_hat))

estimate_mu_2 = sim_results_df |>
  mutate(p.value = ifelse(
    p_value < 0.05, "was rejected", "fail to reject")) |>
  filter(p.value == "was rejected" ) |>
  group_by(mus) |>
  summarise(average_estimate_1 = mean(mu_hat))

plot_q3_new = estimate_mu |>
  ggplot() +
  geom_line(
    aes(x = mus,
        y = average_estimate,
        color = "red")) +
  geom_point(
    aes(x = mus,
        y = average_estimate,
        color = "red")) +
  geom_line(data = estimate_mu_2,
    aes(x = mus,
        y = average_estimate_1,
        color = "yellow")) +
  geom_point(data = estimate_mu_2,
    aes(x = mus,
        y = average_estimate_1,
        color = "yellow")) +
  labs(x = "True Mean",
       y = "Average Estimate",
       title = "True Mean vs Average Estimate")
  

plot_q3_new
```
The sample average of μ^ across tests for which the null is rejected is approximately equal to the true value of μ when the true mean is greater than 4.
The sample average of μ^ across tests for which the null is rejected is not equal to the true value of μ when the true mean is less than 4.
The reason for the above trend is because when the true mean is greater than 4, the power and effect size is very large (close to 1).
